Modal type theory based on the intuitionistic epistemic logic

### 2018-04-22 article.tex
added brand new lemma in article
update in article

\documentclass[a4paper]{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bussproofs}
\usepackage{mathtools}
\usepackage{verbatim}
\usepackage{dsfont}
\usepackage{mathabx}
\usepackage[all, 2cell]{xy}
\usepackage[all]{xy}
\newtheorem{theorem}{Theorem}
\newtheorem{prop}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{defin}{Definition}
\newtheorem{ex}{Example}
\newtheorem{col}{Corollary}
\newtheorem{con}{Consequence}
\usepackage{listings} 		% for source code
\date{}
\title{Modal type theory based on the intuitionistic epistemic logic}

@@

  % This is a comment, not shown in final output.
  % The following shows typesetting  power of LaTeX:

@@

  Modal intutionistic epistemic logic IEL was proposed by S. Artemov and T. Proropopescu \cite{Artemov}. IEL
  provides the epistimology and the theory of knowledge as based on BHK-semantics of intuitionistic logic.
  IEL$^{-}$  is a variant of IEL, that corresponds to intuitionistic belief.
  Informally, ${\bf K}A$ denotes that $A$ is verified intuitionistically.

@@

    2) ${\bf K}(A \to B) \to ({\bf K}A \to {\bf K}B)$ (normality);

    3) $A \to {\bf K}A$ (co-reflection);

@@

  Functional programming languages such as Haskell \cite{Haskell}, Idris \cite{Idris}, Purescript \cite{Purs}
  Elm \cite{Elm} or Scala \cite{Scala} have special type classes\footnote{Type class in Haskell is a general interface for
  special group of datatypes.} for calculations with container types like \verb"Functor" and
  \verb"Applicative" \footnote{Reader may read more about container types in the Haskell standard library documentation\cite{Base} or in the next one textbook
  \cite{Miran}}:

@@

  We investigate the relationship between intuitionistic epistemic logic IEL$^{-}$ and applicative programming with side-effects
  by constructing the type system (which is called $\lambda_{{\bf K}}$) which is Curry-Howard isomorphic to $IEL^{-}$.
  So we will consider ${\bf K}$-modality as an arbitrary applicative functor and we prove that obtained type system
  is sound and complete for applicative functor on cartesian closed category (using the categorical definition proposed by Paterson \cite{Cons}).

  $\lambda_{{\bf K}}$ consists of the rules for simply typed lambda-calculus and special typing rules for
  lifting types into the applicative functor ${\bf K}$. We assume that our type system will
  axiomatize the simplest case of computation with effects with one container. We provide a proof-theoretical
  view at this kind of computations in functional programming and prove strong normalization and confluence.

@@

  \begin{minipage}{0.5\textwidth}
    \begin{flushleft}
  	\begin{prooftree}
    	\AxiomC{$\Gamma \vdash A$}
    	\RightLabel{${\bf K}_I$}
    	\UnaryInfC{$\Gamma \vdash {\bf K}A$}
  \end{prooftree}
    \end{flushleft}
  \end{minipage}
  \begin{minipage}{0.5\textwidth}
    \begin{flushright}
  	\begin{prooftree}
    \AxiomC{$\Gamma \vdash {\bf K} A_1, \dots, \Gamma \vdash {\bf K} A_n $}
    \AxiomC{$A_1,\dots,A_n \vdash B$}
    \BinaryInfC{$\Gamma \vdash {\bf K} B$}
  	\end{prooftree}
    \end{flushright}
  \end{minipage}
  \end{defin}

@@

We will denote $\Gamma \vdash {\bf K} A_1, \dots, \Gamma \vdash {\bf K} A_n$ and $A_1,\dots,A_n \vdash B$ as $\Gamma \vdash {\bf K} \vec{A}$ and $\vec{A} \vdash B$ for brevity.

@@

1) If $\Gamma \vdash_{\text{NIEL}^{-}} A$, then $\text{IEL}^{-} \vdash \bigwedge \Gamma \rightarrow {\bf K}A$.

$\begin{array}{lll}
(1) & \bigwedge \Gamma \rightarrow A & \text{assumption}\\
(2) & A \rightarrow {\bf K}A &\text{co-reflection}\\
(3) & (\bigwedge \Gamma \rightarrow A) \rightarrow ((A \rightarrow {\bf K}A) \rightarrow (\bigwedge \Gamma \rightarrow {\bf K}A))&\text{IPC theorem}\\
(4) & (A \rightarrow {\bf K}A) \rightarrow (\bigwedge \Gamma \rightarrow {\bf K}A) &\text{from (1), (3) and MP}\\
(5) & \bigwedge \Gamma \rightarrow {\bf K}A &\text{from (2), (4) and MP}\\
\end{array}$

@@

2) If $\Gamma \vdash_{\text{NIEL}^{-}} {\bf K} \vec{A}$ and $\vec{A} \vdash B$, then $\text{IEL}^{-} \vdash \bigwedge \Gamma \rightarrow {\bf K}B$.

$\begin{array}{lll}
(1) &\bigwedge \Gamma \rightarrow \bigwedge \limits_{i = 1}^{n} {\bf K} A_i & \text{assumption} \\
(2) &\bigwedge \limits_{i = 1}^{n} {\bf K} A_i \rightarrow {\bf K} \bigwedge \limits_{i = 1}^{n} A_i& \text{IEL$^{-}$ theorem} \\
(3) &\bigwedge \Gamma \rightarrow {\bf K} \bigwedge \limits_{i = 1}^{n} A_i & \text{from (1), (2) and transitivity} \\
(4) &\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B& \text{assumption} \\
(5) &(\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B) \rightarrow {\bf K} (\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B)& \text{co-reflection}\\
(6) &{\bf K} (\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B)& \text{from (4), (5) and MP} \\
(7) &{\bf K} \bigwedge \limits_{i = 1}^{n} A_i \rightarrow {\bf K} B & \text{from (6) and normality} \\
(8) &\bigwedge \Gamma \rightarrow {\bf K} B & \text{from (3), (7) and transitivity}
\end{array}$

@@

  Let $\mathbb{V}$ be the set of variables. The set $\Lambda_{{\bf K}}$ of terms is defined by the grammar:

$\begin{array}{lll}
& \Lambda_{{\bf K}} ::= \mathbb{V} \: | \:  (\lambda \mathbb{V}.\Lambda_{{\bf K}}) \: | \: (\Lambda_{{\bf K}}\Lambda_{{\bf K}}) \: | \: (\Lambda_{{\bf K}} , \Lambda_{{\bf K}}) \: | \: (\pi_1 \Lambda_{{\bf K}}) \: | \: (\pi_2 \Lambda_{{\bf K}}) \: | & \\
& \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad ({\bf pure \: } \: \Lambda_{{\bf K}}) \: | \: ({\bf let \:\: pure} \:\: \mathbb{V}^{*} = \Lambda_{{\bf K}}^{*} \:\: {\bf in} \:\: \Lambda_{{\bf K}})
\end{array}$

@@

Where $\mathbb{V}^{*}$ and $\Lambda_{{\bf K}}^{*}$ denote the set of finite sequences of variables $\bigcup \limits_{i=0}^{\infty} \mathbb{V}^i$
and the set of finite sequences of terms $\bigcup \limits_{i = 0}^{\infty} {\Lambda_{{\bf K}}}^i $. Note that the sequence of variables $\vec{x}$ and the sequence of terms $\vec{M}$ should have the same length.
Otherwise, term is not well-formed.

@@

  Let $\mathbb{T}$ be the set of atomic types. The set $\mathbb{T}_{{\bf K}}$ of types with applicative
  functor ${\bf K}$ is generated by the grammar:
  \begin{equation}
    \mathbb{T}_{{\bf K}} ::= \mathbb{T} \: | \: (\mathbb{T}_{{\bf K}} \to \mathbb{T}_{{\bf K}}) \: |
    \: (\mathbb{T}_{{\bf K}} \times \mathbb{T}_{{\bf K}}) \: | \: ({\bf K}\mathbb{T}_{{\bf K}})
  \end{equation}
  \end{defin}

@@

  \begin{defin} Modal typed lambda calculus $\lambda_{{\bf K}}$ based on NIEL$^{-}_{\land, \to}$:

@@

      \begin{prooftree}
        \AxiomC{$\Gamma \vdash M : A$}
        \RightLabel{${\bf K}_I$}
        \UnaryInfC{$\Gamma \vdash {\bf pure \: } \: M : {\bf K}A $}
      \end{prooftree}
  \end{minipage}%
  \hfill
  \begin{minipage}{0.45\textwidth}
  \begin{tabular}{p{\textwidth}}
    \begin{prooftree}
    \AxiomC{$\Gamma \vdash M : A \to B$}
    \AxiomC{$\Gamma \vdash N : A$}
    \RightLabel{$\rightarrow_e$}
    \BinaryInfC{$\Gamma \vdash MN : B$}
    \end{prooftree}

@@

    \begin{prooftree}
      \AxiomC{$\Gamma \vdash \vec{M} : {\bf K} \vec{A}$}
      \AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
      \RightLabel{$\text{let}_{{\bf K}}$}
      \BinaryInfC{$\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N : {\bf K} B$}
    \end{prooftree}
  \end{tabular}
  \end{minipage}%

@@

  ${\bf K}_I$-typing rule is the same as $\bigcirc$-introduction in lax logic (also known as monadic
  metalanguage \cite{Lax}) and in typed lambda-calculus which is derived by proof-assignment for lax-logic
  proofs.
  ${\bf K}_I$ allows to inject an object of type $\alpha$ into the functor. ${\bf K}_I$  reflects the
  Haskell method {\bf pure} for Applicative class. It plays the same role as the {\bf return} method in Monad class.

  $\text{let}_{{\bf K}}$ is similar to the $\Box$-rule in typed lambda calculus for intuitionistic normal modal logic {\bf IK}, which is described in \cite{ModalK}.

$\Gamma \vdash \vec{M} : {\bf K} \vec{A}$ is a syntax sugar for the sequence $\Gamma \vdash M_1 : {\bf K}A_1,\dots,\Gamma \vdash M_n : {\bf K}A_n$ and $\vec{x} : \vec{A} \vdash N : B$ is a short form for $x_1 : A_1, \dots, x_n : A_n \vdash N : B$.
${\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N$ is a simultaneous local binding in $N$.
We use this short form instead of ${\bf let \: pure\:} x_1,\dots,x_n = M_1,\dots,M_n {\: \bf in \:} N$.

In fact, our calculus is the extention of typed lambda calculus for {\bf IK} by ${\bf K}_I$-rule that is appropriate to co-reflection.

@@

Here are some examples of closed terms:

\begin{prooftree}
\AxiomC{$x : A \vdash x : A$}
\UnaryInfC{$x : A \vdash {\bf pure \:} x : {\bf K} A$}
\UnaryInfC{$\vdash (\lambda x. {\bf pure \: } x) : A \to {\bf K}A$}
\end{prooftree}

\begin{prooftree}
\AxiomC{$f : {\bf K} (A \to B) \vdash f : {\bf K} (A \to B)$}
\AxiomC{$x : {\bf K} A \vdash x : {\bf K} A $}
\AxiomC{$g : A \to B \vdash g : A \to B$}
\AxiomC{$y : A \vdash y : A$}
\RightLabel{$\to_e$}
\BinaryInfC{$g : A \to B, y : A \vdash g y : B$}
\RightLabel{$\text{let}_{{\bf K}}$}
\TrinaryInfC{$f : {\bf K} (A \to B), x : {\bf K} A \vdash {\bf let \: pure \:} g, y  = f, x {\: \bf in \:} g y : {\bf K} B$}
\UnaryInfC{$f : {\bf K} (A \to B) \vdash \lambda x. {\bf let \: pure\:} g, y = f, x {\: \bf in \:} g y : {\bf K} A \to {\bf K} B$}
\UnaryInfC{$\vdash \lambda f. \lambda x. {\bf let \:pure \:} g, y = f, x {\: \bf in \:} g y : {\bf K}(A \to B) \to {\bf K} A \to {\bf K} B$}
\end{prooftree}

@@

  Now we define free variables and substitutions. $\beta$-reduction, multi-step $\beta$-reduction and $\beta$-equality are defined standardly:

@@

  1) $FV(x) = \{ x \}$;

  2) $FV(\lambda x. M) = FV(M) \setminus \{ x\}$;

  3) $FV(M N) = FV(M) \cup FV(N);$

  4) $FV(\langle M,N \rangle) = FV(M) \cup FV(N)$;

  5) $FV(\pi_i M) \subseteq FV(M)$, $i \in \{ 1, 2\}$;

  6) $FV(\text{pure } M) = FV(M)$;

  7) $FV({\bf let \: pure} \: \vec{x} = \vec{M} \:\: {\bf in} \:\: N) = \bigcup \limits_{i = 1}^n FV(M), \text{where $n = |\vec{M}|$}$.
  \end{defin}

@@

  1) $x [x := N] = N$, $x [y := N] = x$;

  2) $(M N) [x := N] = M[x := N] N [x := N]$;

  3) $(\lambda x. M) [x := N] = \lambda x. M [y := N]$, $y \in FV(M)$;

  4) $(M, N)[x := P] = (M[x := P], N [x := P])$;

  5) $(\pi_i M) [x := P] = \pi_i (M[x := P])$, $i \in \{ 1, 2\}$;

  6) $({\bf pure \: } M) [x := P] = {\bf pure \: } (M [x := P])$;

  7) $({\bf let \: pure \:}\vec{x} = \vec{M} {\: \bf in \:} N) [y := P] = {\bf let \: pure \:} \vec{x} = (\vec{M} [y := P]) {\: \bf in \:} N$.
  \end{defin}

Substitution and free variable for terms of kind ${\bf let \: pure}$ are defined similary to \cite{ModalK}.

  \begin{defin} Type substituition

    The substituition of type $C$ for type variable $B$ in type $A$ is defined inductively:

    1) $B [B := C] = B$ and $D [B := C] = D$, if $B \neq D$;

    2) $(A_1 \alpha A_2)[B := C] = (A_1[B := C]) \alpha (A_2[B := C])$, where $\alpha \in \{ \to, \times \}$;

    3) $({\bf K}A)[B := C] = {\bf K}(A [B := C])$;

    4) Let $\Gamma$ be the context, then $\Gamma [B := C] = \{ x : (A[B := C]) \: | \: x : A \in \Gamma \}$.
  \end{defin}

  \begin{defin} $\beta$-reduction and $\eta$-reduction rules for $\lambda_{{\bf K}}$.

  1) $(\lambda x. M) N \rightarrow_{\beta} M [x := N]$;

  2) $\pi_1 \langle M, N \rangle \rightarrow_{\beta} M$;

  3) $\pi_2 \langle M, N \rangle \rightarrow_{\beta} N$;

  4) $\begin{array}{llll}
  & {\bf let \: pure \:} \vec{x}, y, \vec{z} = \vec{M}, {\bf let \: pure \: } \vec{w} = \vec{N} {\: \bf in \: } Q, \vec{P} {\: in \:} R \rightarrow_{\beta} & \\
  & {\bf let \: pure \:} \vec{x}, \vec{w}, \vec{z} = \vec{M}, \vec{N}, \vec{P} {\: \bf in \: } R [y := Q]
  \end{array}$

  5) ${\bf let \: pure \:} \vec{x} = {\bf pure \:} \vec{M} {\: \bf in \:} N \rightarrow_{\beta} {\bf pure \:} N [\vec{x} := \vec{M}]$

  6) ${\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} M \rightarrow_{\beta} {\bf pure \:} M$, where \underline{\quad} is an empty sequence of terms.

  7) $\lambda x. f x \rightarrow_{\eta} f$;

  8) $\langle \pi_1 P, \pi_2 P \rangle \rightarrow_{\eta} P$;

  9) ${\bf let \: pure \:} x = M {\: \bf in \: } x \rightarrow_{\eta} M$;

@@

This commuting relation denotes that we identify terms of kind {\bf let pure} if they are differ in the order of binding.

@@

\begin{lemma} Generation for ${\bf K}_I$.

  Let $\Gamma \vdash {\bf pure \:} M : {\bf K}A$, then $\Gamma \vdash M : A$;
\end{lemma}

@@

\begin{itemize}
  \item If $\Gamma \vdash M : A$ and $\Gamma \subseteq \Delta$, then $\Delta \vdash M : A$;
  \item If $\Gamma \vdash M : A$, then $\Delta \vdash M : A$, where $\Delta = \{ x_i : A_i \: | \: (x_i : A_i) \in \Gamma \: \& \: x_i \in FV(M) \}$
  \item If $\Gamma, x : A \vdash M : B$ and $\Gamma \vdash N : A$, then $\Gamma \vdash M [x := N] : B$.
  \item If $\Gamma \vdash M : A$, then $\Gamma [B := C] \vdash M : (A [B := C])$.
\end{itemize}
\end{lemma}

@@

\begin{prooftree}
  \AxiomC{$\Gamma \vdash \vec{M} : {\bf K} \vec{A}$}
  \AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
  \RightLabel{$\text{let}_{{\bf K}}$}
  \BinaryInfC{$\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N : {\bf K}B$}
\end{prooftree}

By IH $\Delta \vdash \vec{M} : {\bf K} \vec{A}$, so $\Delta \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N : {\bf K}B$.

@@

2) Similary.

\begin{prooftree}
  \AxiomC{$\Gamma, z : C \vdash \vec{M} : {\bf K} \vec{A}$}
  \AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
  \RightLabel{$\text{let}_{{\bf K}}$}
  \BinaryInfC{$\Gamma, z : C \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\: \bf in \: } N : {\bf K}B$}
\end{prooftree}

Let $\Gamma \vdash P : C$.

By IH, $\Gamma \vdash \vec{M} [z := P] : {\bf K} \vec{A}$. So $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \vec{M} [z := P] {\: \bf in \: } N : {\bf K}B$

4-5) Similary.

@@

  If $\Gamma \vdash M : A$ and $M \twoheadrightarrow_{\beta \eta} N$, then $\Gamma \vdash N : A$

@@

Induction on the derivation $\Gamma \vdash M : A$ and on the generation of $\rightarrow_{\beta \eta}$.

For cases with application, abstraction and pairs see \cite{Morten} \cite{Pierce}.

>>>>

\documentclass[a4paper]{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{bussproofs}
\usepackage{mathtools}
\usepackage{verbatim}
\usepackage{dsfont}
\usepackage{mathabx}
\usepackage{stmaryrd}
\usepackage[all, 2cell]{xy}
\usepackage[all]{xy}
\newtheorem{theorem}{Theorem}
\newtheorem{prop}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{defin}{Definition}
\newtheorem{ex}{Example}
\newtheorem{col}{Corollary}
\newtheorem{con}{Consequence}
\usepackage{listings} 		% for source code
\date{}
\title{Modal type theory based on the intuitionistic epistemic logic}

@@

@@

  Modal intutionistic epistemic logic IEL was proposed by S. Artemov and T. Proropopescu \cite{Artemov}. IEL
  provides the epistimology and the theory of knowledge as based on BHK-semantics of intuitionistic logic.
  IEL$^{-}$  is a variant of IEL, that corresponds to intuitionistic belief.
  Informally, $\Box A$ denotes that $A$ is verified intuitionistically.

@@

    2) $\Box (A \to B) \to (\Box A \to \Box B)$ (normality);

    3) $A \to \Box A$ (co-reflection);

@@

  Functional programming languages such as Haskell \cite{Haskell}, Idris \cite{Idris} or Purescript \cite{Purs}
  have special type classes\footnote{Type class in Haskell is a general interface for
  special group of datatypes.} for calculations with container types like \verb"Functor" and
  \verb"Applicative" \footnote{Reader may read more about container types in the Haskell standard library documentation\cite{Base} or in the next one textbook
  \cite{Miran}}:

@@

  We investigate the relationship between intuitionistic epistemic logic IEL$^{-}$ and applicative programming with side-effects
  by constructing the type system (which is called $\lambda_{\text{IEL}^{-}}$) which is Curry-Howard isomorphic to $IEL^{-}$.
  So we will consider IEL$^{-}$ modality as an arbitrary applicative functor and we prove that obtained type system
  is sound and complete for applicative functor on cartesian closed category (using the categorical definition proposed by Paterson \cite{Cons}).

  $\lambda_{\text{IEL}^{-}}$ consists of the rules for simply typed lambda-calculus and special typing rules for
  lifting types into the applicative functor $\Box$. We assume that our type system will
  axiomatize the simplest case of computation with effects with one container. We provide a proof-theoretical
  view at this kind of computations in functional programming and prove strong normalization and confluence.

@@

  \begin{minipage}{0.5\textwidth}
    \begin{flushleft}
  	\begin{prooftree}
    	\AxiomC{$\Gamma \vdash A$}
    	\RightLabel{$\Box_I$}
    	\UnaryInfC{$\Gamma \vdash \Box A$}
  \end{prooftree}
    \end{flushleft}
  \end{minipage}
  \begin{minipage}{0.5\textwidth}
    \begin{flushright}
  	\begin{prooftree}
    \AxiomC{$\Gamma \vdash \Box A_1, \dots, \Gamma \vdash \Box A_n $}
    \AxiomC{$A_1,\dots,A_n \vdash B$}
    \BinaryInfC{$\Gamma \vdash \Box B$}
  	\end{prooftree}
    \end{flushright}
  \end{minipage}
  \end{defin}

@@

We will denote $\Gamma \vdash \Box A_1, \dots, \Gamma \vdash \Box A_n$ and $A_1,\dots,A_n \vdash B$ as $\Gamma \vdash \Box \vec{A}$ and $\vec{A} \vdash B$ for brevity.

@@

1) If $\Gamma \vdash_{\text{NIEL}^{-}} A$, then $\text{IEL}^{-} \vdash \bigwedge \Gamma \rightarrow \Box A$.

$\begin{array}{lll}
(1) & \bigwedge \Gamma \rightarrow A & \text{assumption}\\
(2) & A \rightarrow \Box A &\text{co-reflection}\\
(3) & (\bigwedge \Gamma \rightarrow A) \rightarrow ((A \rightarrow \Box A) \rightarrow (\bigwedge \Gamma \rightarrow \Box A))&\text{IPC theorem}\\
(4) & (A \rightarrow \Box A) \rightarrow (\bigwedge \Gamma \rightarrow \Box A) &\text{from (1), (3) and MP}\\
(5) & \bigwedge \Gamma \rightarrow \Box A &\text{from (2), (4) and MP}\\
\end{array}$

@@

2) If $\Gamma \vdash_{\text{NIEL}^{-}} \Box \vec{A}$ and $\vec{A} \vdash B$, then $\text{IEL}^{-} \vdash \bigwedge \Gamma \rightarrow \Box B$.

$\begin{array}{lll}
(1) &\bigwedge \Gamma \rightarrow \Box A_1, \dots, \bigwedge \Gamma \rightarrow \Box A_1 & \text{assumption} \\
(2) &\bigwedge \Gamma \rightarrow \bigwedge \limits_{i = 1}^{n} \Box A_i & \text{IEL$^{-}$ theorem} \\
(3) &\bigwedge \limits_{i = 1}^{n} \Box A_i \rightarrow \Box \bigwedge \limits_{i = 1}^{n} A_i & \text{IEL$^{-}$ theorem} \\
(4) &\bigwedge \Gamma \rightarrow \Box \bigwedge \limits_{i = 1}^{n} A_i & \text{from (1), (2) and transitivity} \\
(5) &\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B& \text{assumption} \\
(6) &(\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B) \rightarrow \Box (\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B)& \text{co-reflection}\\
(7) &\Box (\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B)& \text{from (4), (5) and MP} \\
(8) &\Box \bigwedge \limits_{i = 1}^{n} A_i \rightarrow \Box B & \text{from (6) and normality} \\
(9) &\bigwedge \Gamma \rightarrow \Box B & \text{from (3), (7) and transitivity}
\end{array}$

@@

  Let $\mathbb{V}$ be the set of variables. The set $\Lambda_{\Box}$ of terms is defined by the grammar:

$\begin{array}{lll}
& \Lambda_{\Box} ::= \mathbb{V} \: | \:  (\lambda \mathbb{V}.\Lambda_{\Box}) \: | \: (\Lambda_{\Box} \Lambda_{\Box}) \: | \: \langle \Lambda_{\Box} , \Lambda_{\Box} \rangle \: | \: (\pi_1 \Lambda_{\Box}) \: | \: (\pi_2 \Lambda_{\Box}) \: | & \\
& \quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad ({\bf box \: } \:\Lambda_{\Box}) \: | \: ({\bf let \: box \:} \mathbb{V}^{*} = \Lambda_{\Box}^{*} \:\: {\bf in} \:\: \Lambda_{\Box})
\end{array}$

@@

Where $\mathbb{V}^{*}$ and $\Lambda_{\Box}^{*}$ denote the set of finite sequences of variables $\bigcup \limits_{i=0}^{\infty} \mathbb{V}^i$
and the set of finite sequences of terms $\bigcup \limits_{i = 0}^{\infty} \Lambda_{\Box}^i $. Note that the sequence of variables $\vec{x}$ and the sequence of terms $\vec{M}$ should have the same length.
Otherwise, term is not well-formed.

@@

  Let $\mathbb{T}$ be the set of atomic types. The set $\mathbb{T}_{\Box}$ of types with applicative
  functor $\Box$ is generated by the grammar:
  \begin{equation}
    \mathbb{T}_{\Box} ::= \mathbb{T} \: | \: (\mathbb{T}_{\Box} \to \mathbb{T}_{\Box} \: |
    \: (\mathbb{T}_{\Box} \times \mathbb{T}_{\Box}) \: | \: (\Box \mathbb{T}_{\Box})
  \end{equation}
  \end{defin}

@@

  \begin{defin} Modal typed lambda calculus $\lambda_{\text{IEL}^{-}}$ based on NIEL$^{-}_{\land, \to}$:

@@

      \begin{prooftree}
        \AxiomC{$\Gamma \vdash M : A$}
        \RightLabel{$\Box_I$}
        \UnaryInfC{$\Gamma \vdash {\bf box \: } \: M : \Box A $}
      \end{prooftree}
  \end{minipage}%
  \hfill
  \begin{minipage}{0.45\textwidth}
  \begin{tabular}{p{\textwidth}}
    \begin{prooftree}
    \AxiomC{$\Gamma \vdash M : A \to B$}
    \AxiomC{$\Gamma \vdash N : A$}
    \RightLabel{$\rightarrow_e$}
    \BinaryInfC{$\Gamma \vdash MN : B$}
    \end{prooftree}

@@

    \begin{prooftree}
      \AxiomC{$\Gamma \vdash \vec{M} : \Box \vec{A}$}
      \AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
      \RightLabel{$\text{let}_{\Box}$}
      \BinaryInfC{$\Gamma \vdash {\bf let \: box \:} \vec{x} = \vec{M} {\: \bf in \: } N : \Box B$}
    \end{prooftree}
  \end{tabular}
  \end{minipage}%

@@

  $\Box_I$-typing rule is the same as $\bigcirc$-introduction in monadic
  metalanguage \cite{Lax}. $\Box_I$ allows to inject an object of type $A$ into the functor $\Box$. $\Box_I$ reflects the
  Haskell method {\bf box} for Applicative class. It plays the same role as the {\bf return} method in Monad class.

  $\text{let}_{\Box}$ is similar to the $\Box$-rule in typed lambda calculus for intuitionistic normal modal logic {\bf IK}, which is described in \cite{ModalK}.

$\Gamma \vdash \vec{M} : \Box \vec{A}$ is a syntax sugar for the sequence $\Gamma \vdash M_1 : \Box A_1,\dots,\Gamma \vdash M_n : \Box A_n$ and $\vec{x} : \vec{A} \vdash N : B$ is a short form for $x_1 : A_1, \dots, x_n : A_n \vdash N : B$.
${\bf let \: box \:} \vec{x} = \vec{M} {\: \bf in \: } N$ is a simultaneous local binding in $N$.
We use this short form instead of ${\bf let \: box \:} x_1,\dots,x_n = M_1,\dots,M_n {\: \bf in \:} N$.

In fact, our calculus is the extention of typed lambda calculus for {\bf IK} by $\Box_I$-rule that is appropriate to co-reflection.

@@

Here are some examples of derivation trees:

\begin{prooftree}
\AxiomC{$x : A \vdash x : A$}
\UnaryInfC{$x : A \vdash {\bf box \:} x : \Box A$}
\UnaryInfC{$\vdash (\lambda x. {\bf box \: } x) : A \to \Box A$}
\end{prooftree}

\begin{prooftree}
\AxiomC{$f : \Box (A \to B) \vdash f : \Box (A \to B)$}
\AxiomC{$x : \Box A \vdash x : \Box A $}
\AxiomC{$g : A \to B \vdash g : A \to B$}
\AxiomC{$y : A \vdash y : A$}
\RightLabel{$\to_e$}
\BinaryInfC{$g : A \to B, y : A \vdash g y : B$}
\RightLabel{$\text{let}_{\Box}$}
\TrinaryInfC{$f : \Box (A \to B), x : \Box A \vdash {\bf let \: box \:} g, y  = f, x {\: \bf in \:} g y : \Box B$}
\UnaryInfC{$f : \Box (A \to B) \vdash \lambda x. {\bf let \: box \:} g, y = f, x {\: \bf in \:} g y : \Box A \to \Box B$}
\UnaryInfC{$\vdash \lambda f. \lambda x. {\bf let \: box \:} g, y = f, x {\: \bf in \:} g y : \Box (A \to B) \to \Box A \to \Box B$}
\end{prooftree}

@@

  Now we define free variables and substitutions:

@@

\begin{enumerate}
  \item $FV(x) = \{ x \}$;
  \item $FV(\lambda x. M) = FV(M) \setminus \{ x\}$;
  \item $FV(M N) = FV(M) \cup FV(N);$
  \item $FV(\langle M,N \rangle) = FV(M) \cup FV(N)$;
  \item $FV(\pi_i M) \subseteq FV(M)$, $i \in \{ 1, 2\}$;
  \item $FV(\text{\bf box } M) = FV(M)$;
  \item $FV({\bf let \: box} \: \vec{x} = \vec{M} {\: \bf in \:} N) = \bigcup \limits_{i = 1}^n FV(M), \text{where $n = |\vec{M}|$}$.
\end{enumerate}
  \end{defin}

@@

\begin{enumerate}
  \item $x [x := N] = N$, $x [y := N] = x$;
  \item $(M N) [x := N] = M[x := N] N [x := N]$;
  \item $(\lambda x. M) [x := N] = \lambda x. M [y := N]$, $y \in FV(M)$;
  \item $(M, N)[x := P] = (M[x := P], N [x := P])$;
  \item $(\pi_i M) [x := P] = \pi_i (M[x := P])$, $i \in \{ 1, 2\}$;
  \item $({\bf box \: } M) [x := P] = {\bf box \: } (M [x := P])$;
  \item $({\bf let \: box \:}\vec{x} = \vec{M} {\: \bf in \:} N) [y := P] = {\bf let \: box \:} \vec{x} = (\vec{M} [y := P]) {\: \bf in \:} N$.
\end{enumerate}
  \end{defin}

Substitution and free variable for terms of kind ${\bf let \: box}$ are defined similary to \cite{ModalK}.

  \begin{defin} $\beta$-reduction and $\eta$-reduction rules for $\lambda_{\text{IEL}^{-}}$.

\begin{enumerate}
  \item $(\lambda x. M) N \rightarrow_{\beta} M [x := N]$;
  \item $\pi_1 \langle M, N \rangle \rightarrow_{\beta} M$;
  \item $\pi_2 \langle M, N \rangle \rightarrow_{\beta} N$;
  \item ${\bf let \: box \:} \vec{x}, y, \vec{z} = \vec{M}, {\bf let \: box \: } \vec{w} = \vec{N} {\: \bf in \: } Q, \vec{P} {\: in \:} R \rightarrow_{\beta} \\
  {\bf let \: box \:} \vec{x}, \vec{w}, \vec{z} = \vec{M}, \vec{N}, \vec{P} {\: \bf in \: } R [y := Q]$
  \item ${\bf let \: box \:} \vec{x} = {\bf box \:} \vec{M} {\: \bf in \:} N \rightarrow_{\beta} {\bf box \:} N [\vec{x} := \vec{M}]$
  \item ${\bf let \: box \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} M \rightarrow_{\beta} {\bf box \:} M$, where \underline{\quad} is an empty sequence of terms.
  \item $\lambda x. f x \rightarrow_{\eta} f$;
  \item $\langle \pi_1 P, \pi_2 P \rangle \rightarrow_{\eta} P$;
  \item ${\bf let \: box \:} x = M {\: \bf in \: } x \rightarrow_{\eta} M$;
\end{enumerate}
  \end{defin}

@@

@@

\begin{lemma} Generation for $\Box_I$.

  Let $\Gamma \vdash {\bf box \:} M : \Box A$, then $\Gamma \vdash M : A$;
\end{lemma}

@@

\begin{enumerate}
  \item If $\Gamma \vdash M : A$ and $\Gamma \subseteq \Delta$, then $\Delta \vdash M : A$;
  \item If $\Gamma \vdash M : A$, then $\Delta \vdash M : A$, where $\Delta = \{ x_i : A_i \: | \: (x_i : A_i) \in \Gamma \: \& \: x_i \in FV(M) \}$
  \item If $\Gamma, x : A \vdash M : B$ and $\Gamma \vdash N : A$, then $\Gamma \vdash M [x := N] : B$.
\end{enumerate}
\end{lemma}

@@

\begin{prooftree}
  \AxiomC{$\Gamma \vdash \vec{M} : \Box \vec{A}$}
  \AxiomC{$\vec{x} : \vec{A} \vdash N : B$}
  \RightLabel{$\text{let}_{\Box}$}
  \BinaryInfC{$\Gamma \vdash {\bf let \: box \:} \vec{x} = \vec{M} {\: \bf in \: } N : \Box B$}
\end{prooftree}

By IH $\Delta \vdash \vec{M} : \Box \vec{A}$, so $\Delta \vdash {\bf let \: box \:} \vec{x} = \vec{M} {\: \bf in \: } N : \Box B$.

@@

2)-3) Similary.

@@

  If $\Gamma \vdash M : A$ and $M \twoheadrightarrow_r N$, then $\Gamma \vdash N : A$

@@

Induction on the derivation $\Gamma \vdash M : A$ and on the generation of $\rightarrow_r$.

For cases with axiom, application, abstraction and pairs see \cite{Morten} \cite{Pierce}.
